# Introduction to Computational Finance

## Optimal Portfolios

In this series we deal with the closing prices of Macdonald’s, Bank of America, IBM,Chevron, Coca-Cola, Novartis and AT&T, over a one-year time span extending from 2013-05-01 to 2014-05-01.
Using the data available on Moodle, draw Markowitz’s efficient frontier using on the one hand the analytical expression, and on the other a random Monte-Carlo search.
Using the efficient frontier, find the weight of the portfolio with the minimal volatility.
What can you say about the return of this portfolio?

## methode_monte_carlo:
```{r}
mafct <- function(data) {
poids<-prop.table(runif(7))
ret<-data[-1,]
for (i in 1:nrow(ret)){
  for (j in 1:ncol(ret)){
    ret[i,j]<-(ret[i,j]-data[i,j])/data[i,j]
  }
}
C <- var(ret,na.rm = T)
esperance_retours<-colMeans(ret)
multip_ret<-replicate(7,1)
for (i in 1:nrow(ret)) {
   multip_ret<-multip_ret*(1+ret[i,])
   }
portfolio_esp_tab<-poids*esperance_retours
portfolio_esp<-sum(portfolio_esp_tab)
portfolio_var<-0
for (i in 1:nrow(C)) {
  for (j in 1:ncol(C)) {
    portfolio_var<-portfolio_var+poids[i]*poids[j]*C[i,j]
  }
}

portfolio_sigma=sqrt(portfolio_var)
temp<- list(mu_p=as.numeric(portfolio_esp),sqrt_var_p=as.numeric(portfolio_sigma))
return (temp)
}

data <- read.table("closes.dat", 
                   col.names = c("mcdo","boa","ibm","ch","coca","nov","att"));
N=1000
xx<-0
yy<-0
for (i in 1:N){
  r<-mafct(data)
  xx<-c(xx,r$sqrt_var_p)
  yy<-c(yy,r$mu_p)
}
xx<-xx[-1]
yy<-yy[-1]
plot(xx, yy,
     main="efficient_frontier_monteCarlo",
     xxlab="sigma_p",
     yylab="E(R_p)")
```

## methode_analytique_plusPlots:
```{r}
data <- read.table("closes.dat", 
                   col.names = c("mcdo","boa","ibm","ch","coca","nov","att"));

mafct2 <- function(data) {
means <- colMeans(data,na.rm = T) #vest 1,7 qui contient les moyennes
C <- var(data,na.rm = T) #matric de cov
invC <- solve(C) #inverse la matrice

un <- matrix(1, nrow = length(means))


alpha <- t(un) %*% invC %*% un #oy t(..) est la transposee
beta <- t(un) %*% invC %*% means
gamma <- t(means) %*% invC %*% means
delta <- alpha * gamma - beta * beta

a = as.numeric(alpha)
b = as.numeric(beta)
c = as.numeric(gamma)
d = as.numeric(delta)

weights <- prop.table((runif(7)))
mu_p<-sum(weights*means)

lambd1<-(c-b*mu_p)/d
lambd2<-(a*mu_p-b)/d

w <- invC%*%t((lambd1*t(un)+lambd2*means))
sigma_p_sq <- t(w)%*%C%*%w
#test<-lambd1+lambd2*mu_p
temp<- list(mu_p=as.numeric(mu_p),var_p=as.numeric(sigma_p_sq),poids=c(weights))
return (temp)
}
ret<-data[-1,]
for (i in 1:nrow(ret)){
  for (j in 1:ncol(ret)){
    ret[i,j]<-(ret[i,j]-data[i,j])/data[i,j]
  }
}
N=10000
x<-0
y<-0
w<-prop.table((runif(7)))
for (i in 1:N){
  r<-mafct2(ret)
  x<-c(x,r$var_p)
  y<-c(y,r$mu_p)
  w<-c(w,r$poids)
}
x<-x[-1]
y<-y[-1]
w<-w[-1]
w<-w[-1]
w<-w[-1]
w<-w[-1]
w<-w[-1]
w<-w[-1]
w<-w[-1]

plot(x, y,
     main="efficient_frontier_avecmin",
     xlab="sigma_p",
     ylab="E(R_p)",
     #ylim = c(-1e-04 ,max(y))
)

#-------
index_min_sigma_p<-which.min(x)
t<-index_min_sigma_p*7
poids_corresp<-c(w[t-6],w[t-5],w[t-4],w[t-3],w[t-2],w[t-1],w[t])
print(poids_corresp)
points(x[index_min_sigma_p],y[index_min_sigma_p],col='red')
#----------
plot(xx, yy,
     main="efficient_frontier",
     xlab="sigma_p",
     ylab="E(R_p)",
     #ylim = c(-1e-04 ,max(y))
     )
points(sqrt(x), y,col='red')
legend("topright",
       c("ef mithode analytique","mithode Monte Carlo"),
       fill=c("red","black"),
       cex=0.6)
```

## return-computation
```{r}

mafct <- function(data) {#je mets un seul poids ici et je veux calculer le return de ce portfolio
  poids<-c(0.16615240, 0.17895094, 0.30915982, 0.18169583, 0.05915160, 0.01160653, 0.09328287)
  
  ret<-data[-1,]
  for (i in 1:nrow(ret)){
    for (j in 1:ncol(ret)){
      ret[i,j]<-(ret[i,j]-data[i,j])/data[i,j]
    }
  }
  
  C <- var(ret,na.rm = T)
  
  esperance_retours<-colMeans(ret)
  
  multip_ret<-replicate(7,1)
  for (i in 1:nrow(ret)) {
    multip_ret<-multip_ret*(1+ret[i,])
  }
  multip_ret<-multip_ret*poids
  
  portfolio_esp_tab<-poids*esperance_retours
  portfolio_esp<-sum(portfolio_esp_tab)
  
  #print(portfolio_esp)
  
  portfolio_var<-0
  for (i in 1:nrow(C)) {
    for (j in 1:ncol(C)) {
      portfolio_var<-portfolio_var+poids[i]*poids[j]*C[i,j]
    }
  }
  
  portfolio_sigma=sqrt(portfolio_var)
  #print(portfolio_sigma)
  temp<- list(mu_p=as.numeric(portfolio_esp),sqrt_var_p=as.numeric(portfolio_sigma),retour_p=sum(multip_ret))
  return (temp)
}

data <- read.table("closes.dat", 
                   col.names = c("mcdo","boa","ibm","ch","coca","nov","att"));

tempp<-mafct(data)
print(tempp$retour_p)
debut<-sum(data[1,])


N=1000
xx<-0
yy<-0
for (i in 1:N){
  r<-mafct(data)
  xx<-c(xx,r$sqrt_var_p)
  yy<-c(yy,r$mu_p)
}

xx<-xx[-1]
yy<-yy[-1]

plot(xx, yy,
     main="efficient_frontier_monteCarlo",
     xxlab="sigma_p",
     yylab="E(R_p)")

```

* a)
On the graph "efficient_frontier", the Markowitz's efficient frontier has been calculated in red using using the analytical method. In black are the pairs obtained with the Monte Carlo method. 
Carlo method.  I did my code for 10,000 iterations for the analytical solution. But I think I have a But I think I have an error in my Monte Carlo code because it takes a long time to run. I have only only run it for 1,000 iterations, but if you run more iterations for the Monte Carlo method you will see a very close result between the two methods.

* b)
If you are risk averse, for example, you will want a minimum variance. In our case, we want to minimise sigma_p (standard deviation). The point in red in this second diagram The red point in this second diagram, "efficient_frontier_avecmin", corresponds to the minimum of sigma_p. In my example, the red point corresponds to the couple (3.66e-05,-3.25e-05) with the weights : (rounded values)
Macdonald's Bank of America IBM Chevron Coca-Cola Novartis AT&T
0.17 0.18 0.31 0.18 0.06 0.01 0.09
with a portfolio return of about 0.98. So you lose money in this case if you are risk averse. Note: if we run this example again, we will have different value weights but a return very close to the one found above. return very close to the one found above