# Introduction to Computational Finance

## Time average
Draw the simple moving average for N = 100 and the exponential moving averages for
α = 2/(N + 1), α = 0.1 and α = 0.01.

```{r}
set.seed(0);
x=100+cumsum(0.5-runif(1000))
N=100
#MA
MA<-c(1:N)
temp=0
for (i in 1:N){
  temp=temp+x[i+N]
}
MA[1]=temp/N

for (i in 2:N){
  MA[i]=MA[i-1]-(x[i-1]/N)+(x[i+N-1]/N)
}
#EMA
EMA1<-c(1:N)
alpha1=0.1

temp1=0
temp2=0
for (i in 1:N){
  temp1=temp1+x[i+N]*(1-alpha1)^(i-1)
  temp2=temp2+(1-alpha1)^(i-1)
}
EMA1[1]=temp1/temp2
for (i in 2:N){
  EMA1[i]=alpha1*x[i+N]+(1-alpha1)*EMA1[i-1]
}
#-----
EMA2<-c(1:N)
alpha2=0.01
temp2_1=0
temp2_2=0
for (i in 1:N){
  temp2_1=temp2_1+x[i+N]*(1-alpha2)^(i-1)
  temp2_2=temp2_2+(1-alpha2)^(i-1)
}
EMA2[1]=temp2_1/temp2_2
for (i in 2:N){
  EMA2[i]=alpha2*x[i+N]+(1-alpha2)*EMA2[i-1]
}
#----
EMA3<-c(1:N)
alpha3=2/(N+1)
temp3_1=0
temp3_2=0
for (i in 1:N){
  temp3_1=temp3_1+x[i+N]*(1-alpha3)^(i-1)
  temp3_2=temp3_2+(1-alpha3)^(i-1)
}
EMA3[1]=temp3_1/temp3_2
for (i in 2:N){
  EMA3[i]=alpha3*x[i+N]+(1-alpha3)*EMA3[i-1]
}

plot(EMA1,typ='l', col="blue",
     main="Exponential_moving_average",
     xlab="temps",
     ylab="valeur")
lines(EMA2,typ='l',col="green")
lines(EMA3,typ='l',col="red")
lines(MA,typ='l',col="brown")
legend("topleft",
       c("alpha=0.1","alpha=0.01","alpha=2/(N+1)","MA"),
       fill=c("blue","green","red","brown"),
       cex=0.6)
```

* What can be said regarding the case α = 0.1 compared to α = 0.01 ?
If we take an alpha close to 1, this will give us Zt.
So if we take an alpha that is 0.1, this will reduce the oldest observations more
the oldest observations more quickly, whereas an alpha of 0.01 will
alpha of 0.01 will take the older values into account for longer.
older values for longer.

* If we take alpha=2/(N+1), this will depend on N. Thus, if N
is 100 as in the beginning of the exercise, we will have alpha
which is about 0.02, and therefore reduces its observations slightly more
observations compared to alpha worth 0.1.


## Scaling law

Using the data provided (EUR/USD, tick-by-tick, from January 1st to
March 1st 2012), compute the number of directional changes for different scales δ.
Draw a log-log plot of the number of directional changes as a function of the scale δ.
What do you observe ?

```{r}
data <- read.table("EUR_USD-20120101-20120301.dat", col.names = c("timestamp","bid","ask"));

data$timestamp <- as.POSIXct(data$timestamp, origin="1970-01-01")

b <- seq(as.POSIXct("2012-01-01 00:00:00"),as.POSIXct("2012-03-01 00:00:00"), by = "15 min")

data22 <- data[data$timestamp >= "2012-01-01",]

data22$mid<-(data22$bid+data22$ask)/2
#----

algo<- function(delta,data2){
modeup=1#ie initialisi ` up
#delta<-0.1
xext<-data2$mid[1]
nbChange=0
for (x in data2$mid){
  if (modeup==1){
    if (x<xext){
      xext<-x
    }
    else if (((x-xext)/xext)>=delta){
      #print("new DC up")
      xext<-x
      modeup=0
      nbChange=nbChange+1
    }
  }
  else{
    if (x>xext){
      xext<-x
    }
    else if (((xext-x)/xext)>=delta){
      #print("new DC down")
      xext<-x
      modeup=1
      nbChange=nbChange+1
    }
  }
}
return(nbChange)}

ww<-seq(0.01, 0.0001, -0.0001)
a<-c(1:100)
incr=0.01
for (i in 1:100){
  a[i]<-algo(incr,data22)
  incr=incr-0.0001
}

plot(a,col="blue",typ='l',
     main = "Nombre_de_chgmts_de_directions",
     log="xy",
     xlab="valeur de l'intervalle",
     ylab="nombre de chgmt de directions")
```

It can be seen that the number of changes of direction
direction increases exponentially as the interval is reduced.
the interval. This confirms what we have seen in class with the idea of a fractal curve.
seen in class with the idea of a fractal curve.
Note: on the graph
Number_of_chgmts_of_directions, on the x-axis the
first value is 0.01 and at each step, we decrease this value
this value of 0.0001