# TP1 Introduction to Computational Finance

## Computing returns

The following series describes the value of a portfolio at the end of each month:  
  
Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec  
101 102 103 104 105 106 107 108 109 110 111 112  
  
The initial value of the portfolio being 100, the annual return is 12%.

* Compute monthly returns
```{r}
x<-seq(101,112)
monthlyReturn<-rep(0,12)

for (i in 2:12){monthlyReturn[i]<-(x[i]-x[i-1])/x[i-1]}
monthlyReturn[1]<-(101-100)/100
print(format(round(monthlyReturn, 4), nsmall = 4))
```

* Compute the annual return from the monthly returns. Compare it to the sum of the
monthly returns. Which calculation leads to the correct result, i.e. annual return of
12%.

```{r}
sumOfMonthlyReturn<-0
for (i in 1:12){sumOfMonthlyReturn<-sumOfMonthlyReturn+monthlyReturn[i]}
cat('sum of the monthly returns:', sumOfMonthlyReturn, "\n")
annRetFromMonthlyRet<-100
for(i in 1:12){annRetFromMonthlyRet=annRetFromMonthlyRet*(1+monthlyReturn[i])}
annRetFromMonthlyRet=(annRetFromMonthlyRet-100)/100
cat('annual return from the monthly returns:', annRetFromMonthlyRet)
```

* Compute the average monthly return and the average of the monthly returns
```{r}
averageMonthlyReturn<-((112-100)/100)/12
cat('average monthly return:', averageMonthlyReturn, "\n")
averageOfTheMonthlyRet <-0
for (i in 1:12){averageOfTheMonthlyRet <- averageOfTheMonthlyRet + monthlyReturn[i]}
averageOfTheMonthlyRet <- averageOfTheMonthlyRet/12
cat('average of the monthly returns:', averageOfTheMonthlyRet)
```
## Portfolio of Microsoft and Starbucks stock

Consider the case where you purchase 10 shares of each Microsoft and Starbucks stock at
the end of month t − 1 at prices $P_{msft,t−1} = \$85$ and $P_{sbux,t−1} = \$30$.

* Compute the initial value of the portfolio.
```{r}
initialPorteFolio<-10*85+10*30
cat('initial value of the portfolio:',initialPorteFolio)
```

* Compute the portfolio shares $α_{msft}$ and $α_{sbux}$.
```{r}
alphaMsf=10/20
alphaSbux=10/20
cat('alpha_{msft}:',alphaMsf,'\n')
cat('alpha_{sbux}:',alphaSbux)

```
Consider now that at the end of month t, the prices are $P_{msft,t−1} = \$90$ and $P_{sbux,t−1} = \$28$.  

* Compute the one-period return of Microsoft and Stabuck stocks.
```{r}
returnMsf=(90-85)/85
returnSbux=(28-30)/30
cat('return of Microsoft:',returnMsf,'\n')
cat('return of Starbuck:',returnSbux)
```

* Finally, compute the return of the portfolio and its value at the end of month t.
```{r}
returnPortefolio=((10*90+10*28)-initialPorteFolio)/initialPorteFolio
finalPortefolio=10*90+10*28
cat('return of the portfolio:',returnPortefolio,'\n')
cat('portefolio value at the end of month t:',finalPortefolio)
```
